---
title: "Mississipi Exploratory Work"
author: "Ellen Graham"
date: "12/4/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Some packages
```{r}
library(tidyverse)
library(readr)
library(sf)
library(glue)
library(lubridate)
library(rnaturalearthdata)
library(rnaturalearth)
```


## Load in the data!
```{r, warning=FALSE}
# all warnings come from empty line on end of files and can be ignored
leg_0_col_names <- c("index", 
               "crew", 
               "date", 
               "time_GMT", 
               "latitude", 
               "longitude",
               "alt",
               "speed", 
               "heading",
               "climb",
               "accel_x",
               "accel_y", 
               "accel_z",
               "gyro_x",
               "gyro_y",
               "gyro_z", 
               "air_temp",
               "air_gas",
               "air_humid",
               "air_pressure",
               "water_temp", 
               "ysi_ph",
               "ysi_do",
               "ysi_no3",
               "ysi_sal",
               "ysi_tur",
               "flow")

leg_0_col_types <- list(col_number(),
                  col_character(),
                  col_date(format = "%Y-%m-%d"),
                  col_time(format = "%H:%M:%S"),
                  col_number(), 
                  col_number(), 
                  col_number(),
                  col_number(),
                  col_number(),
                  col_number(),
                  col_number(),
                  col_number(),
                  col_number(),
                  col_number(),
                  col_number(),
                  col_number(),
                  col_number(),
                  col_number(),
                  col_number(),
                  col_number(),
                  col_number(),
                  col_number(),
                  col_number(),
                  col_number(),
                  col_number(),
                  col_number(),
                  col_number())

leg_2_col_names <- c(leg_0_col_names, "comp_time")
leg_2_col_types <- rlist::list.append(leg_0_col_types, col_integer())

leg_0_file_names <- list("raw_data/2019830-datalog.txt",
                         "raw_data/2019831-datalog.txt",
                         "raw_data/20190902-datalog.txt")

leg_2_file_names <- setdiff(list.files(path = "raw_data", 
                               pattern = ".txt",
                               full.names = TRUE),
                            leg_0_file_names)

leg_0_data <- map(leg_0_file_names, read_csv,
                  col_names = leg_0_col_names,
                  col_types = leg_0_col_types,
                  na = c("None")) %>% 
  bind_rows()

leg_2_data <- map(leg_2_file_names, read_csv,
                  col_names = leg_2_col_names,
                  col_types = leg_2_col_types,
                  na = c("None", ""),
                  skip_empty_rows = TRUE) %>% 
  bind_rows()

```

Some notes: Leg 2 water temp not accurate, crew list not accurate


## Cleaning up the data:
- Removing index, ysi_ph, ysi_d0, ysi_no3, ysi_sal, ysi_tur, flow from both leg 1 and leg 2 as they are 0 or NA for all observations
- Set crew list and water temp to NA for leg 2
- Adding combined datetime
- Adding proper index to each observation

```{r}
leg_0_data <- leg_0_data %>% 
  select(-ysi_ph, 
         -ysi_do, 
         -ysi_no3, 
         -ysi_sal, 
         -ysi_tur, 
         -flow, 
         -index) %>% 
  mutate(leg = 0)

leg_2_data <- leg_2_data %>% 
  select(-ysi_ph, 
         -ysi_do, 
         -ysi_no3, 
         -ysi_sal, 
         -ysi_tur, 
         -flow,
         -index,
         -comp_time) %>% 
  mutate(leg = 2) %>% 
  mutate(crew = NA, water_temp = NA)




all_legs <- bind_rows(leg_0_data, leg_2_data) %>% 
  mutate(date_time = parse_datetime(paste(date, time_GMT), 
                                    format = "%Y-%m-%d %H:%M:%S")) %>% 
  group_by(leg) %>% 
  mutate(index = 1:n()) %>% 
  ungroup() %>% 
  select(index, crew, date, time_GMT, date_time, everything())
#warnings expected, come from NAs in date and time
```


I'm going to remove all observations without a time associated with them. There are 6000 cases where time is out of order already, so I feel trying to interpolate based on the time is inapporpriate, as we have no guarentee those times are correct.

I'm also going to remove observations where air_gas and related measurements weren't taken, as there are only 16 total and only 2 after removing missing times. 


```{r}
all_legs_small <- all_legs %>% 
  filter(!is.na(date)) %>% 
  arrange(leg,date_time) %>% 
  filter(!is.na(air_pressure))
```



## Exploring

Make line graph for all numeric vars to see if any obvious errors in the data
```{r, eval=FALSE}
num_vars <- all_legs_small %>% 
  select_if(is.numeric) %>% 
  select(-index) %>% 
  names()


produce_point_plot <- function(y_name, data_set) {
  g <- ggplot(data_set) +
    geom_point(aes_string(x = "index", y = y_name)) +
    labs(title = glue("{y_name} over observations"))
  return(g)
}

map(num_vars, produce_point_plot, all_legs_small)
```

Looks like a lot of sensors give 0 when it obviously shouldn't be. 
I'm going to set 0's for air_temp, air_gas, air_humid, air_pressure to NA

To help fix spikes in data, I'm going to take the median value in each 15? second chunk of time for all points
```{r}
all_legs_med <- all_legs_small %>% 
  mutate(air_temp = ifelse(air_temp == 0, NA, air_temp),
         air_gas = ifelse(air_gas == 0, NA, air_gas),
         air_humid = ifelse(air_humid == 0, NA, air_humid),
         air_pressure = ifelse(air_pressure == 0, NA, air_pressure)) %>% 
  mutate(time_in_sec = as.numeric(date_time)) %>% 
  mutate(fifteen_sec_chunk = (time_in_sec%/%15)*15) %>% 
  group_by(fifteen_sec_chunk) %>% 
  summarise_at(6:23, median, na.rm = TRUE) %>% 
  mutate(date_time = as_datetime(fifteen_sec_chunk)) %>% 
  mutate(index = 1:n())
 

num_vars_med <- all_legs_med %>% 
  select_if(is.numeric) %>% 
  names()

# map(num_vars_med, produce_point_plot, all_legs_med)
 
```

I'm pretty happy with the results, lets start building maps!

First a couple more steps, lets make lat+long into a simple features geometry object

```{r}
clean_legs <- st_as_sf(all_legs_med, coords = c("longitude", "latitude"), crs = 4326)
```

Load up a mississippi river file
```{r}
river_50 <- st_as_sf(ne_download(scale = 10, type = 'rivers_lake_centerlines', category = 'physical'))

miss_basin <- st_read("data\\miss_basin\\Miss_RiverBasin.shp") %>% 
  st_transform(crs = 4326)

```


```{r}
us_sf <- ne_states(country = "united states of america", returnclass = "sf")

tiny_legs = clean_legs[seq(1, nrow(clean_legs), 100), ]


miss_rivers <- river_50 %>% 
  filter(name %in% c("Mississippi", "Ohio", "Missouri"))
states_in_miss_basin <- st_intersection(miss_basin, us_sf)
ggplot() + 
  #geom_sf(data = miss_basin, fill = "#D3DBC6") +
  geom_sf(data = states_in_miss_basin, fill = "#FAFAFA") +
  geom_sf(data = miss_rivers, color = "#409CFF") +
  geom_sf(data = tiny_legs, aes(color = heading)) +
  coord_sf(xlim = c(-115.00, -77.00), ylim = c(27.00, 50.00), expand = FALSE, datum = NA) +
  theme_minimal() +
  scale_color_viridis_c(option = "plasma")

```


Download clean legs
```{r, eval=FALSE}
readr::write_csv(all_legs_med, "clean_data\\clean_data_non_geom.csv")
BRRR::skrrrahh(0)
```

